Home Nighttime Cough Monitoring

### 2025-03-05 Lilah Melzer/cough_monitor_analysis.qmd
added baseline survey table summaries

```{r}
zip1 <- mixed_model(Coughs ~ Night + Diagnosis,
  data = agg_counts,
  random = ~ 1 | sid, zi_random = ~ 1 | sid,
  family = zi.poisson(), zi_fixed = ~1
)
zip2 <- update(zip1, zi_fixed = ~ Night + Diagnosis)

@@

So, instead of the model approach I did a more simple correlation between median coughs per night and median respiratory score.

### CFQ-R Caregivers 6-13

#### Median scores

```{r}
#| warning: false
cfqr_parent_6_13$sid <- as.numeric(cfqr_parent_6_13$sid)
caregivers <-
  full_join(agg_counts, cfqr_parent_6_13, join_by(sid, Date)) %>%
  select(sid, Sick, Night, Coughs, respirat) %>%
  filter(!if_all(Coughs:respirat, ~ is.na(.)))
caregivers_med <- caregivers %>%
  group_by(sid) %>%
  summarise(
    `Median Coughs` = median(Coughs, na.rm = T),
    `Median Resp. Score` = median(respirat, na.rm = T)
  ) %>%
  drop_na()
M <- cor.test(caregivers_med$`Median Coughs`,
  caregivers_med$`Median Resp. Score`,
  method = "spearman"
)
ggplot(caregivers_med, aes(
  y = `Median Coughs`, x = `Median Resp. Score`,
  colour = factor(sid)
)) +
  geom_point() +
  theme_bw()
```

Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).

#### Plots by participant

```{r}
#| message: false
#| warning: false
invisible(lapply(unique(caregivers_med$sid), function(s) {
  d <- caregivers[caregivers$sid == s, ]
  d <- d %>% pivot_longer(c(Coughs, respirat))
  p <- ggplot(d, aes(x = Night, y = value, color = Sick)) +
    geom_point() +
    geom_line() +
    theme_bw() +
    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
    facet_wrap(~name, ncol = 1) +
    ggtitle(paste0("Participant ", s))
  cat("\n")
  cat("\n")
  print(p)
  cat("\n")
  cat("\n")
}))
```

### CFQ-R Children Ages 12-13

#### Median scores

```{r}
#| warning: false
cfqr_children_12_13$sid <- as.numeric(cfqr_children_12_13$sid)
children_12_13 <-
  full_join(agg_counts, cfqr_children_12_13, join_by(sid, Date)) %>%
  select(sid, Sick, Night, Coughs, respirat) %>%
  filter(!if_all(Coughs:respirat, ~ is.na(.)))
children_12_13_med <- children_12_13 %>%
  group_by(sid) %>%
  summarise(
    `Median Coughs` = median(Coughs, na.rm = T),
    `Median Resp. Score` = median(respirat, na.rm = T)
  ) %>%
  drop_na()
M <- cor.test(children_12_13_med$`Median Coughs`,
  children_12_13_med$`Median Resp. Score`,
  method = "spearman"
)
ggplot(children_12_13_med, aes(
  y = `Median Coughs`, x = `Median Resp. Score`,
  colour = factor(sid)
)) +
  geom_point() +
  theme_bw()
```

Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).

#### Plots by participant

```{r}
#| message: false
#| warning: false
invisible(lapply(unique(children_12_13_med$sid), function(s) {
  d <- children_12_13[children_12_13$sid == s, ]
  d <- d %>% pivot_longer(c(Coughs, respirat))
  p <- ggplot(d, aes(x = Night, y = value, color = Sick)) +
    geom_point() +
    geom_line() +
    theme_bw() +
    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
    facet_wrap(~name, ncol = 1) +
    ggtitle(paste0("Participant ", s))
  cat("\n")
  cat("\n")
  print(p)
  cat("\n")
  cat("\n")
}))
```

### CFQ-R Adolescents and Adults 14 Years

#### Median scores

```{r}
#| warning: false
cfqr_teens_adults$sid <- as.numeric(cfqr_teens_adults$sid)
teens_adults <-
  full_join(agg_counts, cfqr_teens_adults, join_by(sid, Date)) %>%
  select(sid, Sick, Night, Coughs, respirat) %>%
  filter(!if_all(Coughs:respirat, ~ is.na(.)))
teens_adults_med <- teens_adults %>%
  group_by(sid) %>%
  summarise(
    `Median Coughs` = median(Coughs, na.rm = T),
    `Median Resp. Score` = median(respirat, na.rm = T)
  ) %>%
  drop_na()
M <- cor.test(teens_adults_med$`Median Coughs`,
  teens_adults_med$`Median Resp. Score`,
  method = "spearman"
)
ggplot(teens_adults_med, aes(
  y = `Median Coughs`, x = `Median Resp. Score`,
  colour = factor(sid)
)) +
  geom_point() +
  theme_bw()
```

Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).

#### Plots by participant

```{r}
#| message: false
#| warning: false
invisible(lapply(unique(teens_adults_med$sid), function(s) {
  d <- teens_adults[teens_adults$sid == s, ]
  d <- d %>% pivot_longer(c(Coughs, respirat))
  p <- ggplot(d, aes(x = Night, y = value, color = Sick)) +
    geom_point() +
    geom_line() +
    theme_bw() +
    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
    facet_wrap(~name, ncol = 1) +
    ggtitle(paste0("Participant ", s))
  cat("\n")
  cat("\n")
  print(p)
  cat("\n")
  cat("\n")
}))
```

### Healthy controls

#### Median scores

```{r}
#| warning: false
hc_surveys$sid <- as.numeric(hc_surveys$sid)
hcs <-
  full_join(agg_counts, hc_surveys, join_by(sid, Date)) %>%
  select(sid, Sick, Night, Coughs, q1:q5) %>%
  filter(!if_all(q1:q5, ~ is.na(.)))
hcs_med <- hcs %>%
  group_by(sid) %>%
  summarise(
    `Median Coughs` = median(Coughs, na.rm = T),
    `Median Resp. Score` = median(as.numeric(q1), na.rm = T)
  ) %>%
  drop_na()
M <- cor.test(hcs_med$`Median Coughs`,
  hcs_med$`Median Resp. Score`,
  method = "spearman"
)
ggplot(hcs_med, aes(
  y = `Median Coughs`, x = `Median Resp. Score`,
  colour = factor(sid)
)) +
  geom_point() +
  theme_bw()
```

Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).

>>>>

```{r}
# Change reference group for easier interpretation
# agg_counts$Diagnosis <- relevel(agg_counts$Diagnosis, ref = "CF")
# Model selection
zip1 <- mixed_model(Coughs ~ Night + Diagnosis,
  data = agg_counts,
  random = ~ 1 | sid, zi_random = ~ 1 | sid,
  family = zi.poisson(), zi_fixed = ~1
)
zip2 <- update(zip1, zi_fixed = ~ Night + Diagnosis)

@@

A total of `r length(unique(cfqr_children_6_11$sid))` people filled out the CFQ-R for children 6-11, with `r length(unique(cfqr_children_6_11$sid[duplicated(cfqr_children_6_11$sid)]))` filling it out more than once.

A total of `r length(unique(cfqr_children_12_13$sid))` people filled out the CFQ-R for children 12-13, with `r length(unique(cfqr_children_12_13$sid[duplicated(cfqr_children_12_13$sid)]))` filling it out more than once.

A total of `r length(unique(cfqr_teens_adults$sid))` people filled out the CFQ-R for teens and adults, with `r length(unique(cfqr_teens_adults$sid[duplicated(cfqr_teens_adults$sid)]))` filling it out more than once.

A total of `r length(unique(cfqr_parent_6_13$sid))` people filled out the CFQ-R for teens and adults, with `r length(unique(cfqr_parent_6_13$sid[duplicated(cfqr_parent_6_13$sid)]))` filling it out more than once.

## Summary of baseline values

```{r results='asis'}
# Combine CFQ-R into one
cf1 <- cfqr_children_12_13 %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  select(-Date) %>%
  pivot_longer(physical:digest)
cf2 <- cfqr_children_6_11 %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  select(-Date) %>%
  pivot_longer(physical:digest)
cf3 <- cfqr_teens_adults %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  select(-Date) %>%
  pivot_longer(physical:digest)
cf4 <- cfqr_parent_6_13 %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  select(-Date) %>%
  pivot_longer(physical:weight)
# Put together then back to wide
cfqr_baseline <- do.call(rbind, list(cf1, cf2, cf3, cf4))
cfqr_baseline <- cfqr_baseline %>%
  pivot_wider()
t3 <- tableby(
  Survey ~ notest(physical) + notest(emotion) + notest(social) + notest(eat) +
    notest(body) + notest(treat) + notest(respirat) + notest(digest) +
    notest(role) + notest(vitality) + notest(health) + notest(weight) +
    notest(school),
  data = cfqr_baseline, control = mycontrols
)
summary(t3)
```

# Healthy control surveys

A total of `r length(unique(hc_surveys$sid))` controls filled out the surveys, with `r length(unique(hc_surveys$sid[duplicated(hc_surveys$sid)]))` filling it out more than once.

## Summary of baseline values

```{r results='asis'}
hc <- hc_surveys %>%
  group_by(sid) %>%
  filter(row_number() == 1)
t4 <- tableby(
  ~ notest(q1) + notest(q2) + notest(q3) + notest(q4) +
    notest(q5),
  data = hc, control = mycontrols
)
summary(t4)
```
